---
title: "Use Animal Tracks"
author: "John Kirwan"
date: "30/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Get data from animal tracks

```{r message=FALSE, warning=FALSE}
library('tidyverse')
library('magrittr')
library('circular')
library('RColorBrewer')
```

Uses stimulus centre and point opposite stimulus centre as markers for the unit circle on arena. Gets complete track from each set of trials, makes a plot and determines headings.

```{r}
folder <- 'tracks/'
filez  <- list.files(normalizePath(folder),recursive=TRUE)
file_tbl <- tibble(basename=sub("*_point01.txt","",#subset string before exp
             filez[grep('*point01.txt$',filez)]))
file_tbl %<>% mutate(landmark=paste0(basename,"_point01.txt"),
                    track    =paste0(basename,"_point02.txt"))
tail(file_tbl)
```

Define the radius of the inner circle.

**NB:** As the tracks are uncalibrated, the size of objects in the central region are expanded relative to those at the periphery. This should become much more eccentric at the very edges and effect little change near the centre. The inner and outer radii are therefore somewhat smaller than the stated values. 

The radius criterion decides whether the first point with a radius larger than the specified inner and outer radii is used or whether the nearest point either side of the radius is used.

```{r Basic criteria}
circ0 <- .25#.25#.2 # proportion of inner radius covered by inner pt
circ1 <- .5 #.5#.4 # proportion of inner radius covered by outer pt
radius_criterion   <- "first_past"     #"nearest"
stim_mid_criterion <- "dark_trough"#"mid_inflection" 
```

Make functions to bring angle values back into the fold

```{r}
rad2clock   <- function(theta){theta<- -theta+(pi/2)#theta + (pi/2)#
                if(theta < 0){theta <- theta + 2*pi
                }else{ if(theta > 2*pi){theta <- theta - 2*pi} }
                return(theta)}

rad_unclock <- function(theta){theta<- -theta+(pi/2)#theta + (pi/2)#
                if(theta < -pi){theta <- theta + 2*pi
                }else{ if(theta > pi){theta <- theta - 2*pi} }
                return(theta)}

rad_shepherd <- function(theta){  if(theta < -pi){theta <- theta + 2*pi
} else{ if(theta > pi){theta <- theta - 2*pi} }
  return(theta)}

rad_shepherd_clock <- function(theta){  if(theta < 0){theta <- theta + 2*pi
} else{ if(theta > 2*pi){theta <- theta - 2*pi} }
  return(theta)}
```


#### Bring in data frame with video names and trial order

Obst means obstructed - there was initially an obstacle in the way.

```{r}
df <- read.delim('Diadema_tracks.txt', sep="\t", header = TRUE)
df %<>% filter(locale=="Tenerife")
df %<>% filter(arc %in% c(0,17,40)) #use only conditions in paper
if(any(table(df$video) != 4)==FALSE){print("All remaining sets have 4 trials")}
```
The values of arc in the spreadsheet represent the zero mean crossing for the DoG Stimulus, so we change these here to represent the wavelet (period) width. Modify some conditions and show the first rows.

```{r}
df$locale   <- NULL
df$species  <- NULL
df$predator <- NULL
df$arc[df$arc==40 & df$type=="DoG"]    <- 69
df$arc[df$arc==17 & df$type=="DoG"]    <- 29
df$type[df$type=="grey control"]       <- "control"
df$pattern[df$pattern=="grey control"] <- "control"
# Add "success
df$succ4 <- rep(FALSE,dim(df)[1])
df$succ5 <- rep(FALSE,dim(df)[1])
df$succ6 <- rep(FALSE,dim(df)[1])
df$succ4[df$heading_rad<=45*pi/180|df$heading_rad>315*pi/180]<-TRUE
df$succ5[df$heading_rad<=36*pi/180|df$heading_rad>324*pi/180]<-TRUE
df$succ6[df$heading_rad<=30*pi/180|df$heading_rad>330*pi/180]<-TRUE
df$heading_rad <- mapply(rad_unclock, df$heading_rad) # make radians on unit circle
df %<>% unite("cond", type:arc, remove = FALSE, sep=" ")
df %<>% unite("vid_ord", c(video,pos),remove = FALSE, sep="_")
head(df)
```
Are there any (predicted) track names not represented in the file list? i.e. landmark files without a matching track file?

```{r}
if(any(!file_tbl$track %in% filez)){
  stop(print(paste0("The following track file is missing:",
    file_tbl$track[which(!file_tbl$track %in% filez)])))   }
```

Are any track files in the file list without a match in the landmark files?

```{r}
if(any(!filez[grep('*point02.txt$',filez)] %in% file_tbl$track)){
 stop(print(paste0("The following track file lacks a counterpart: ",
   filez[!filez[grep('*point02.txt$',filez)] %in% file_tbl$track] )))  }
```

Are the tracking files all represented in the data spreadsheet?

```{r}
if(any(!file_tbl$basename %in% df$video)){warning("Extra file(s) removed")
  file_tbl <- file_tbl[file_tbl$basename %in% df$video,]}
#are all videos listed in spreadsheet represented in the dir of tracking files
if(any(!df$file %in% file_tbl$basename)){warning("Missing file(s)!")}
head(file_tbl)
```

Add part above to put trial, cond, type, arc, obstacle into file_tbl and therefore to trackz_tbl and trialz.

```{r}
# which track rows shares vid name with df
file_tbl$arc       <- as.character(rep("",dim(file_tbl)[1]))
file_tbl$type       <- as.character(rep("",dim(file_tbl)[1]))
file_tbl$trial_set  <- as.character(rep("",dim(file_tbl)[1]))

i=1
while(i <= dim(df)[1]){
 file_tbl$arc [which(file_tbl$basename==df$video[i])]      <- df$arc[i]
 file_tbl$type[which(file_tbl$basename==df$video[i])]      <- df$type[i]
 file_tbl$trial_set[which(file_tbl$basename==df$video[i])] <- df$trial_set[i]
 i=i+1    }

file_tbl   %<>% unite("cond",c(type,arc),remove=FALSE, sep="_")
```


Try with first file to get the track

```{r}
animal_track = read.delim(normalizePath(paste0(folder,'/',file_tbl$track[1])),
                          sep="\t", header = FALSE)[,1:3]
names(animal_track) <- c("frame","x","y")
head(animal_track)
```

Try with first file to get the landmarks

```{r}
landmarks = read.delim(normalizePath(paste0(folder,'/',file_tbl$landmark[1])),
                          sep="\t", header = FALSE)[,1:3]
names(landmarks) <- c("frame","x","y")
head(landmarks)

if(length(landmarks$x) %% 2 == 1){ warning("Odd number of landmarks!") }
num_trials = length(landmarks$x)/2 
```

#### The files 

With the assignment pipe *%<>%*, the variable is assigned to itself at the end.The radius of the circle based on the stimulus distance is *r_px*. atan2 outputs angles in radians according to the unit cicle from -pi to pi.

The y values of the landmark and tracks have been reversed to run upwards rather than downwards.

*stim_x_0* and *stim_y_0* are cartesian coords which have there location translated by subtracting the origin and the length scaled to the distance from the origin (and points transformed accordingly). Quadrants 1 to 4 are from the top to the right moving 90d clockwise each position.

```{r}
get_trial_limits <- function(landmarks,animal_track,file_tbl,vid_num){ #,

 num_trials = dim(landmarks)[1]/2 # count number of trials
 # pick out start frame and centre and stimulus coords
 tl <- as_tibble(array(0,dim=c(num_trials,21),
    dimnames = list(NULL, c("first_f","orig_x","orig_y","stim_x","stim_y",
    "last_f","stim_x_0","stim_y_0","stim_theta","r_px","quadrant","video",
    "order","vid_ord","x0","y0","x1","y1","arc","type","trial_set"))), #
    .name_repair="unique")
  tl$video <- as.character(tl$video)
  landmarks$y <- - landmarks$y # invert y axis to read from bottom to top
  
  i = 1 
 while (i <= num_trials){ # for each trial
  tl$first_f[i]<- as.numeric(landmarks[(i*2)-1,1])# trial i start frame
  tl$orig_x[i] <- as.numeric(landmarks[(i*2)-1,2])# trial i centre pt x,y
  tl$orig_y[i] <- as.numeric(landmarks[(i*2)-1,3])# trial i centre pt x,y
  tl$stim_x[i] <- as.numeric(landmarks[i*2,2])# trial i stim centre x,y
  tl$stim_y[i] <- as.numeric(landmarks[i*2,3])# trial i stim centre x,y
       if (i < num_trials){
        tl$last_f[i] <- as.numeric(landmarks[((i+1)*2)-1,1]-1)# trial i last frame
     } else { tl$last_f[i]  <- max(animal_track[,1]) } # max along dimension 1 

 tl$video[i]     <- file_tbl[[vid_num,1]]
 tl$arc[i]       <- file_tbl$arc[file_tbl$basename==tl$video[i]]
 tl$type[i]      <- file_tbl$type[file_tbl$basename==tl$video[i]]
 
 #### modify to account for more trial sets per video 
 tl$trial_set[i] <- file_tbl$trial_set[file_tbl$basename==tl$video[i]]
   i = i + 1  }

 tl$stim_x_0 <- tl$stim_x - tl$orig_x # stimulus cart coords - orig
 tl$stim_y_0 <- tl$stim_y - tl$orig_y
 tl$stim_theta <- atan2(tl$stim_y_0,tl$stim_x_0) 
 tl$r_px       <- sqrt((tl$stim_x_0)^2+(tl$stim_y_0)^2) # pixel radius
 # if the stimulus centre is set to the midpoint, shift all stimulus positions accordingly
 if(stim_mid_criterion=="mid_inflection" & tl$type[1]=="Herm"){#shift .5T clockwise
   # subtracting angles only works in +ve space
   stim_theta    <- mapply(rad2clock, tl$stim_theta) # change to clock
   stim_theta    <- stim_theta - tl$arc*(pi/360) # add T/2 to shift clockwise 
   stim_theta    <- mapply(rad_shepherd_clock,stim_theta) #herd into 0 to 2*pi
   tl$stim_theta <- mapply(rad_unclock,stim_theta) # back to unit circle
 }
 #tl$stim_x_0 <- tl$stim_x_0 / tl$r_px # standardize by the radius 
 #tl$stim_y_0 <- tl$stim_y_0 / tl$r_px # standardize by the radius 
 tl$stim_x_0  <- cos(tl$stim_theta)
 tl$stim_y_0  <- sin(tl$stim_theta)
 
 i=1
 while (i <= num_trials){ 
  if (tl$stim_theta[i] > .25*pi & tl$stim_theta[i] <= .75*pi){
      tl$quadrant[i] <- 1 #"top" #if stimulus in first quadrant
  }else if (tl$stim_theta[i]>= -.25*pi & tl$stim_theta[i] < .25*pi){
      tl$quadrant[i] <- 2#"left"
  }else if (tl$stim_theta[i] >= -.75*pi & tl$stim_theta[i] < -.25*pi ){
      tl$quadrant[i] <- 3#"bottom"
  }else {
      tl$quadrant[i] <-4 } #"right"
  i = i + 1
 }

 return(tl)
}
```

#find each trial and get data
# (i) frame no (ii - v) coords for centre and stimulus (vi) last frame #
# (vii - viii) normalised stimulus centre (ix) theta in rads (anticlock,
# zero=0) (x) rho in pixels 

Get limits from first video 

```{r}
trial_limits <- get_trial_limits(landmarks,animal_track,file_tbl,1) #
```

#### Get animal tracks in polar terms relative to the stimulus

Divide the animal tracks up by trial using the first landmark. Initialize 
translated and scaled to origin (ts0_) tracks for later.

```{r Get animal pos fx}
get_animal_pos <- function(trial_limits,animal_track){
  num_trials = dim(trial_limits)[1] # count number of trials
  
  animal_track$y <- -animal_track$y # invert y so that runs from bottom up
  # make nested tibble of the sets
  trackz <- purrr::map(vector(length=num_trials), tibble)

 i <- 1 # initialize i  
 while (i <= dim(trial_limits)[1]){    # for each trial in a video 

   animal_track$trial_num[ animal_track$frame >= trial_limits$first_f[i] & 
     animal_track$frame <= trial_limits$last_f[i]] <- i # get the trial no for each obs
   XY <- as_tibble(animal_track[animal_track$frame >= trial_limits$first_f[i] & 
        animal_track$frame <= trial_limits$last_f[i],]) # make a tibble of that trial
   XY %<>% add_column(ts0_x=0,ts0_y=0,abs_theta=0,rho=0,video=NA,
                      order=NA,vid_ord=NA,rel_theta=0)
   XY[,5:6] <- sweep(XY[,2:3],2,as.numeric(trial_limits[i,2:3]),"-") # subtract origin
   XY[,5:6] <- XY[,5:6] / trial_limits$r_px[i]
   
   # get theta from Cartesian coordinates
   abs_theta    <- mapply(atan2,XY$ts0_y,XY$ts0_x)
   # reorient abs theta to be clockwise and with a zero at the top
   XY$abs_theta <- mapply(rad2clock,abs_theta)

   # looks like this line below should apply to rel_theta
   XY$rho <- unlist(sqrt(XY[,5]^2 + XY[,6]^2)) #/as.numeric(trial_limits$r_px[i]) #normalised rho
   XY$video <- trial_limits$video[i]
   XY$order <- as.character(i) # order of trials in video

   XY %<>% unite("vid_ord", video:order,remove = FALSE, sep="_")
   stim_theta <- rad2clock(trial_limits$stim_theta[i])
   # reorient stim_theta and get relative position
   XY$rel_theta <- XY$abs_theta - stim_theta
   XY$rel_theta <- mapply(rad_shepherd_clock,XY$rel_theta) # herd back into clock
   XY$rel_theta <- mapply(rad_unclock,XY$rel_theta)# get back between -pi and pi
   # get absolute theta values back too
   XY$abs_theta <- mapply(rad_unclock,XY$abs_theta) # return to unit circle
   XY$rel_x     <- XY$rho*cos(XY$rel_theta)
   XY$rel_y     <- XY$rho*sin(XY$rel_theta)
   trackz[[i]]  <- XY
   # assign tracks less than 40% of arena to be empty
   if(min(XY$rho>.4)){
    warning(paste('Track starts too far from centre:'),XY$video[1])}
   if(any(XY$rho> 1)){
    warning(paste('Track exceeds arena width: '),XY$video[1])}
   
   trial_limits$order[i]   <- as.character(i) 
   trial_limits$vid_ord[i] <- paste0(trial_limits[i,12],"_",i)

  if(radius_criterion=="first_past"){ # if taking 1st point outside radii
  if(!any(XY$rho > circ0)){warning('NO PTS BEYOND INNER CIRC!!')
  } else if(any(XY$rho<circ0) & # if points either side of circles but not between
        any(XY$rho>circ1) & !any(XY$rho>circ0 & XY$rho<circ1)){
  trial_limits$x0[i] <- last(XY$ts0_x[XY$rho<circ0]) #get last pts
  trial_limits$y0[i] <- last(XY$ts0_y[XY$rho<circ0]) #in inner circ
  } else{
  trial_limits$x0[i] <- XY$ts0_x[XY$rho >= circ0][1] #get 1st pts 
  trial_limits$y0[i] <- XY$ts0_y[XY$rho >= circ0][1] #beyond inner circ
  }
  if(any(XY$rho <= circ1) & any(XY$rho > circ1)){ # if pts both sides of outer circ
    trial_limits$x1[i] <- XY$ts0_x[XY$rho >= circ1][1] #first pts beyond
    trial_limits$y1[i] <- XY$ts0_y[XY$rho >= circ1][1] # outer circle
  }else if(!any(XY$rho > circ1)){
    warning(paste("NO POINT BEYOND OUTER RADIUS!!!:",trial_limits$vid_ord[i]))
    trial_limits$x1[i] <- XY$ts0_x[which.max(XY$rho)] # get most distant point
    trial_limits$y1[i] <- XY$ts0_y[which.max(XY$rho)] # get most distant point

  }else{warning(paste("NO POINT WITHIN OUTER RADIUS!!!",trial_limits$vid_ord[i]))
    trial_limits$x1[i] <- XY$ts0_x[i+1] # take the 2nd points outside
    trial_limits$y1[i] <- XY$ts0_y[i+1] # as x0,y0 will be 1st
  }
  }else{
    trial_limits$x0[i] <- XY$ts0_x[which.min(abs(XY$rho - circ0))]
    trial_limits$y0[i] <- XY$ts0_y[which.min(abs(XY$rho - circ0))]
    trial_limits$x1[i] <- XY$ts0_x[which.min(abs(XY$rho - circ1))] #.5
    trial_limits$y1[i] <- XY$ts0_y[which.min(abs(XY$rho - circ1))]
  }
   i = i + 1
 }
 return(list(trackz,trial_limits)) # norm to radius
}
```

Get the all the tracks from a single file. Print one set.

```{r}
trial_track_list   <- get_animal_pos(trial_limits,animal_track)
trackz             <- trial_track_list[[1]]
tail(trackz[[1]])
```

```{r}
trial_limits      <- trial_track_list[[2]]
tail(trial_limits)
```

Now apply this to all files.Then transform the coordinates from all tibbles in the big list.

```{r}
get_all_animal_pos <- function(landmarks,animal_track,file_tbl){
  i <- 1 # for all files in file_tbl
  while(i <= dim(file_tbl)[1]){
    animal_track = read.delim(normalizePath(paste0(folder,'/',file_tbl$track[i])),
                            sep="\t", header = FALSE)[,1:3]
    names(animal_track) <- c("frame","x","y")
    landmarks <- read.delim(normalizePath(paste0(folder,'/',
                      file_tbl$landmark[i])),sep="\t", header = FALSE)[,1:3]
    names(landmarks) <- c("frame","x","y")
    if(length(landmarks$x) %% 2 == 1){
      warning(paste("Odd number of landmarks:",file_tbl$basename[i])) }
    trial_limits <- get_trial_limits(landmarks,animal_track,file_tbl,i) #
    trial_track_list <- get_animal_pos(trial_limits,animal_track) # run get_animal_pos
     if(i == 1){  
       all_trackz       <- trial_track_list[[1]] # populate with first list
       trialz <- trial_track_list[[2]] # first instance from above
     } else {
       all_trackz       <- c(all_trackz,trial_track_list[[1]])
       trialz <- rbind(trialz,trial_track_list[[2]])
    }
    i <- i + 1
  }
 trackz_n_limits <- list(all_trackz,trialz) 
 return(trackz_n_limits)
}
```


```{r}
trackz_n_limits  <- get_all_animal_pos(landmarks,animal_track,file_tbl)
all_trackz       <- trackz_n_limits[[1]]
tail(all_trackz[[1]])
```

```{r}
trialz <- trackz_n_limits[[2]]
tail(trialz)
```

#### Find cases where the inner and outer radii don't differ

None of thse files are in the spreadsheet list and they can be removed. 

```{r}
print(paste(dim(trialz)[1],"trials in limits list pre culling"))
idx <- which(trialz$x0==trialz$x1 & trialz$y0==trialz$y1)
if(length(idx) > 0){warning(paste("Trial",idx,"has errors!"))
  all_trackz[idx] <- NULL
  trialz <- trialz[-c(idx),] }
print(paste(length(all_trackz),"tracks post culling"))
print(paste(dim(trialz)[1],"trials in limits list post culling"))
```

This leaves 10 extra trials not represented in the spreadsheet (df) file.

#### Make tibble for tracks rather than list

```{r}
trackz_tbl <- map_dfr(all_trackz, ~ .x)
```

#### Line-Circle Intercept.

Works when vector intersects circle.

```{r}
linecirc <- function(x0,y0,x1,y1){
  h = 0;k=0;radius=1 # circle origin x,y
  a = (x1 - x0)^2 + (y1 - y0)^2 # vector length
  b = (2*(x1 - x0)*(x0 - h)) + (2*(y1 - y0)*(y0 - k)) # 
  c = (x0 - h)^2 + (y0 - k)^2 - radius^2
  D <- (b^2) - (4*a*c) # discriminant - must be positive
  if(a <= 0){warning("Vector length is not positive!")}
  if(D <= 0){warning(paste0(
    "Discriminant is not positive with x0=",x0," y0=",y0," x1=",x1," y1=",y1))}
  if(c >= 0){warning("Starting pt is outside radius!")}
  
  # the roots for t may be found in the usual quadratic form:
  t <- c(0,0)
  t[1] = (-b + sqrt(D)) / (2*a)
  t[2] = (-b - sqrt(D)) / (2*a) 
  # 1st and 2nd rows of ints correspond to 1st and 2nd root respectively
  ints <- rbind(unlist(c( t[1]*(x1 - x0) + x0, t[1]*(y1 - y0) + y0)),
                unlist(c( t[2]*(x1 - x0) + x0, t[2]*(y1 - y0) + y0)))
   
  d <- c(sqrt(ints[1,1]^2 + ints[1,2]^2),sqrt(ints[2,1]^2 + ints[2,2]^2))
  if(all.equal(d[1],d[2])==FALSE){warning("Intersect is off unit circle!")}
  # 1st and 2nd rows of dist2track correspond to 1st and 2nd root respectively
  dist2track    <- c(0,0) 
  dist2track[1] <- sqrt( (ints[1,1]-x1)^2 + (ints[1,2]-y1)^2 )
  dist2track[2] <- sqrt( (ints[2,1]-x1)^2 + (ints[2,2]-y1)^2 )
  xy <- ints[which.min(dist2track),]
  
  # commented because NAs crash this part
  #if(t[1] < 1 | t[2] > 0){ #which.min(dist2track)
  #  warning('Roots imply vector is weird') }
  return(tibble::tibble(x_int = xy[1],y_int = xy[2])) # return closest to outer pt
}
```

#### Get the headings from intersection with the unit circle.

Use atan2 to get the headings in radians. Then, having changed the class to circular, set them so that 0 refers to the stimulus midpoint and the rotation is clockwise. 

The value *degs* differs from *radians* in beginning at the top of the circle and continuing clockwise from 0 to 360. 

```{r}
intz <- pmap_dfr(list(trialz$x0,trialz$y0,trialz$x1,trialz$y1), linecirc)

trialz$x_int    <- intz$x_int
trialz$y_int    <- intz$y_int

trialz$abs_rads <- mapply(atan2,trialz$y_int,trialz$x_int)
abs_rads_clock  <- mapply(rad2clock,trialz$abs_rads) # convert to be 0 to 2*pi, clockwise   
clockstim       <- mapply(rad2clock,trialz$stim_theta) # stim theta to clock
rel_radians     <- mapply(rad_shepherd_clock,abs_rads_clock - clockstim) #herd in clock
trialz$radians  <- mapply(rad_unclock,rel_radians) # back to unit circle

# degree clockwise variable for the trials and tracks
trialz$degs     <- (180/pi)*mapply(rad2clock,trialz$radians)
trackz_tbl$degs <- (180/pi)*mapply(rad2clock,trackz_tbl$rel_theta)
```

#### Make variables for *successful* orientation to target sector 

Then, plug these into the tracks table. 

```{r}
trialz$succ4 <- rep(FALSE,dim(trialz)[1])
trialz$succ5 <- rep(FALSE,dim(trialz)[1])
trialz$succ6 <- rep(FALSE,dim(trialz)[1])
trialz$succ4[trialz$degs<=45|trialz$degs>315]<-TRUE
trialz$succ5[trialz$degs<=36|trialz$degs>324]<-TRUE
trialz$succ6[trialz$degs<=30|trialz$degs>330]<-TRUE

trackz_tbl$succ4 <- rep(FALSE,dim(trackz_tbl)[1])
trackz_tbl$succ5 <- rep(FALSE,dim(trackz_tbl)[1])
trackz_tbl$succ6 <- rep(FALSE,dim(trackz_tbl)[1])

i = 1
while(i <= dim(trialz)[1]){
  trackz_tbl$succ4[trackz_tbl$vid_ord==trialz$vid_ord[i]] <- trialz$succ4[i]
  trackz_tbl$succ5[trackz_tbl$vid_ord==trialz$vid_ord[i]] <- trialz$succ5[i]
  trackz_tbl$succ6[trackz_tbl$vid_ord==trialz$vid_ord[i]] <- trialz$succ6[i]
  i=i+1
}
```

Temporary fix - remove observations with rho value greater than 1. 

```{r}
trackz_tbl <- trackz_tbl[trackz_tbl$rho <= 1,]
```

Check which trails are also in df

```{r}
# which track rows shares vid name with df
trackz_tbl$arc       <- rep(0,dim(trackz_tbl)[1])
trackz_tbl$type      <- as.character(rep("",dim(trackz_tbl)[1]))
trackz_tbl$trial_set <- as.character(rep("",dim(trackz_tbl)[1]))

i=1
while(i <= dim(df)[1]){
 trackz_tbl$arc [which(trackz_tbl$video==df$video[i])] <- df$arc[i]
 trackz_tbl$type[which(trackz_tbl$video==df$video[i])] <- df$type[i]
 trackz_tbl$trial_set[which(trackz_tbl$video==df$video[i])] <- df$trial_set[i]
  i=i+1
}

trackz_tbl %<>% unite("cond",c(type,arc),remove=FALSE, sep=" ")
trialz     %<>% unite("cond",c(type,arc),remove=FALSE, sep=" ")
```


> Sets for which there were not four completed trials (e.g. due to a loss of motivation) were excluded from analysis.

```{r}
i=1
w <- unique(trialz$trial_set) # new variable as we are deleting rows from trials
while(i <= length(w)){
  if(dim(trialz[trialz$trial_set==w[i],])[1] < 4){
    warning(paste('Fewer than 4 trials in set:',w[i]))
    trialz$radians[trialz$trial_set==w[i]]           <- NA
    trackz_tbl$rel_theta[trackz_tbl$trial_set==w[i]] <- NA
  } else {
  if(dim(trialz[trialz$trial_set==w[i],])[1] > 4){
    warning(paste('More than 4 trials in set:',w[i])) }
  }   
  i=i+1
}
```
Remove those trials entirely

```{r}
trialz <- trialz[!is.na(trialz$radians),]
trackz_tbl <- trackz_tbl[!is.na(trackz_tbl$rel_theta),]
```

Are the tracking files all represented in the data spreadsheet?

```{r}
dim(trialz)[1]==dim(df)[1]
unique(trackz_tbl$trial_set)==unique(trialz$trial_set)
unique(df$trial_set)==unique(trialz$trial_set)
unique(trackz_tbl$trial_set)==unique(trialz$trial_set)
```




#### Make variables for inner radius relative to the stimulus

Make something close to stimulus relative versions of the inner circle intersects. Make angles +ve real and subtract the stimulus to get the relative angular position of the inner circle intersect. Then, turn into Cartesian coordinates.

```{r}
clockINtheta    <- mapply(rad2clock,atan2(trialz$y0,trialz$x0))
clockstim       <- mapply(rad2clock,trialz$stim_theta)
rel_circ0_theta <- mapply(rad_shepherd_clock,clockINtheta-clockstim)
rel_circ0_theta <- mapply(rad_unclock,rel_circ0_theta) 
trialz$rho0     <- sqrt(trialz$x0^2 + trialz$y0^2)
trialz$rel_x0   <- trialz$rho0*cos(rel_circ0_theta)
trialz$rel_y0   <- trialz$rho0*sin(rel_circ0_theta)
```

Make degree class circular

```{r}
trialz$degs <- as.circular(trialz$degs,type="angles",
    units="degrees",zero=pi/2,rotation="clock",template= 'none',modulo='asis')
```


#### Identify tracks

Bring over relevant information from df dataframe to other dataframes

```{r}
trackz_tbl     <- subset(trackz_tbl, arc %in% c(0,17,29,40,69))
trialz         <- subset(trialz, arc %in% c(0,17,29,40,69))
```

Make the arc width a factor

```{r}
trialz$arc       <- as.factor(trialz$arc)
trackz_tbl$arc   <- as.factor(trackz_tbl$arc)
```

Make the arc width a factor. This doesn't seem to influence the plotting order and may be redundant.

#### Remove redundant trials

Only the first four valid trials should be used. Check that the positions are correct.

```{r}
trialz %>% group_by(trial_set) %>% dplyr::distinct(quadrant, .keep_all = TRUE) -> trialz
table(trialz$trial_set)
```

Excise the tracks which don't match the trials

```{r}
trackz_tbl <- trackz_tbl[trackz_tbl$vid_ord %in% trialz$vid_ord,]
```

#### Plotting

*Points are plotting correctly although the zero and rotation are according to the unit circle because the points themselves are oriented that way, e.g. an upward pointing track has a heading of c.1.6.*

For some reason it appears to be using the conventional unit circle zero and rotation (points concentrated at x=1,y=0).

Later, use scale_colour_brewer(palette = "Paired")

#### Absolute position of tracks with regard to the arena (not the stimulus)

```{r abs headings fig, fig.height = 8, fig.width = 7}
how_many <- 150
colz <- c("1"="blue","2"="green","3"="pink","4"="orange")
# geom_path is already facing up, without (pi/2)- , geom_point() is not
trackz_tbl %>% #filter(obstacle=="obst") %>% 
 ggplot() + coord_fixed() +
 geom_path(aes(x=ts0_x,y=ts0_y,group=vid_ord,color=trial_set),
             size=.5)  +
 geom_point(aes(x=ts0_x, y=ts0_y,group=vid_ord,color=trial_set),
             size=.5) + 
  facet_wrap(vars(arc,type),ncol=2) + # obstacle
  annotate("path", color="orange",
  x=0+1*cos(seq(0,2*pi,length.out=100)),y=0+1*sin(seq(0,2*pi,length.out=100))) +
  annotate("path", color="blue",linetype="dashed",
  x=0+circ1*cos(seq(0,2*pi,length.out=100)),
  y=0+circ1*sin(seq(0,2*pi,length.out=100))) +
  annotate("path", color="black",linetype="dashed",
  x=0+circ0*cos(seq(0,2*pi,length.out=100)),
  y=0+circ0*sin(seq(0,2*pi,length.out=100))) +
 geom_point(data=trialz,#[trialz$obstacle=="obst" & as.numeric(trialz$id) <= how_many,],   
  aes(x=x_int,y=y_int,color=trial_set),size=6,shape=1) + #,color=id #,group=cond
  #geom_text(data=trialz,#[trialz$obstacle=="obst" & as.numeric(trialz$id) <= how_many,],
  # aes(x=x_int,y=y_int,color=id,label=id),alpha=1,size=3) +
 geom_segment(data=trialz,#[trialz$obstacle=="obst" & as.numeric(trialz$id) <= how_many,],
   aes(x=x0,y=y0,xend=x_int,yend=y_int,color=trial_set),linetype="dotted") + #,colour=id
  xlab('') + ylab('') + theme_classic() + 
  ggtitle('Plot absolute tracks') + 
  theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + theme(legend.position = "none") + 
   theme(plot.title=element_text(size=14,hjust=0.5), 
        strip.background=element_blank(),strip.text=element_blank())
```

#### Are animals attracted to the top-right (the opening)?

From the plots, there may be attraction in some instances - check with null hypothesis tests.

```{r}
trialz %>% group_by(type,arc) %>%
  summarize( 
    n    = n(),
    ids  = n_distinct(vid_ord),
    mean = as.numeric(round(mean.circular(abs_rads),2)),
    R    = round(rho.circular(abs_rads),2),
    Ray.p= round(rayleigh.test(abs_rads)[[2]],2),
    sd   = round(sd.circular(abs_rads),2)
             ) -> abs_summary
abs_summary
```

#### Relative position of tracks with regard to the stimulus

The trial limits radians and the tracks now align. 

```{r}
RhoCircularRad <- function(x) {
   n <- length(x)   
   sinr <- sum(sin(x))
   cosr <- sum(cos(x))
   result <- sqrt(sinr^2 + cosr^2)/n
   return(result)
}

makerad <- function(x){x <- as.circular(x,type="angles",units="radians", 
  template="none",modulo="asis",zero=0,rotation="clock")
  }
```


### Relative tracks - summary table of conditions

Add variable for successfully falling into target sextant.

```{r Summarize relative headings}
trialz %>% group_by(type,arc) %>%
  summarize( 
    n    = n(),
    mu_d = as.numeric(round(mean.circular(degs))),
    R    = round(rho.circular(degs),2),
    Ray.p= round(rayleigh.test(radians)[[2]],3),
    V.p  = round(rayleigh.test(radians,mu=circular(pi/2))[[2]],3),  #circular(pi/2)
    quad= round(sum(succ4)/length(succ4),2),
    Binp4= round(binom.test(sum(succ4),length(succ4),p=1/4)[[3]],2),
    quint= round(sum(succ5)/length(succ5),2),
    Binp5= round(binom.test(sum(succ5),length(succ5),p=1/5)[[3]],2),
    sext = round(sum(succ6)/length(succ6),2),
    Binp6= round(binom.test(sum(succ6),length(succ6),p=1/6)[[3]],2),
    sd   = round(sd.circular(degs),2),
    mu_r = mean.circular(radians),
             ) -> headings_summ
headings_summ
```

A4 size less margins is 9.54 (11.69-1.4-.75) x 7 (8.27-.75-.52) in.

```{r rel headings fig, fig.height = 7.5, fig.width = 7}
how_many <- 50
linetypez <- c("FALSE"="dashed","TRUE"="solid")
# geom_path is already facing up, without (pi/2)- , geom_point() is not
trackz_tbl %>%  #filter(obstacle=="obst") %>% 
  #filter(between(as.numeric(id),1,how_many)) %>% # view few
 ggplot() + coord_fixed() + 
 geom_path(aes(x=rel_x,y=rel_y, group=vid_ord,color=succ4))+
 geom_point(aes(x=rel_x,y=rel_y,group=vid_ord,color=succ4),size=.3)+ #,size=succ5>0
 facet_wrap(vars(arc,type),ncol=2,strip.position="top") +
 annotate("path", color="orange",
   x=cos(seq(0,2*pi,length.out=100)),y=sin(seq(0,2*pi,length.out=100))) +
 annotate("path", color="blue",linetype="dashed",
   x=circ1*cos(seq(0,2*pi,length.out=100)),
   y=circ1*sin(seq(0,2*pi,length.out=100))) +
 annotate("path", color="black",linetype="dashed",
   x=circ0*cos(seq(0,2*pi,length.out=100)),
   y=circ0*sin(seq(0,2*pi,length.out=100))) +
 geom_point(data=trialz,#[ # Stick on final relative headings
   aes(x=cos(radians),y=sin(radians),color=succ4),size=5,shape=1)+ 
 geom_segment(data=trialz,
   aes(x=rel_x0,y=rel_y0,xend=cos(radians),yend=sin(radians),
     colour=succ4),linetype="dotted") +
 geom_text(data=headings_summ, #panel labels
   aes(x=-.9,y=1,label=paste0(type," ",arc,"°")),size=3.5)+
 geom_segment(data=headings_summ,aes(
   x=0,y=0,xend=R*cos(mu_r),yend=R*sin(mu_r),group=arc),
   size=1.3,colour="black",arrow=arrow(length=unit(.2,"cm")))+
 geom_point(data=df, # spreadsheet headings
   aes(x=1*cos(heading_rad),y=1*sin(heading_rad),colour=succ4),shape=3) +
   xlab('') + ylab('') + theme_classic() + 
   labs(title='Plot tracks relative to stimulus') +
   theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + theme(legend.position = "none") +
   theme(plot.title=element_text(size=14,hjust=0.5), 
        strip.background=element_blank(),strip.text=element_blank())
```

Uncalibrated tracks from the *D. africanum* adult trials. Each figure panel, represents the experiment arena looking downwards from above. The outer orange circle represents the wall of the arena, with is coloured grey, except for a visual stimulus, graded across the horizon according to a 1st Hermitian wavelet. The darkest part of the stimulus is located at the topmost point of the panel. Within the arena are marked tracks (297 in total), beginning near the centre, which delineate the paths taken by individual animals in trials. In these trials, animals must first negotiate an obstruction at the centre. Each colour corresponds to a given individual (of 50 in total). The points just after the track first crosses each of the two (dashed black and blue) concentric circles near the centre are used to mark a vector (shown with a straight line, dashed for headings directed to the target) to the arena wall. The intersection with the wall (the animal's heading) is shown by a hollow circle and the id of the animal. Animals are oriented in the 45d condition (and to a lesser extent the 60d and 150d condition) indicating that this species can see at a resolution better than 45d but worse than 30d.

#### Effect of differing sector sizes

```{r Plot success with differing sectors}
headings_summ %>% pivot_longer(c(quad,quint,sext)) %>% filter(type!="bar") %>%
  ggplot(aes(x=arc,y=value,color=name,group=name)) +
  geom_segment(x=0,xend=150,y=.2,yend=.2,color="darkgreen",linetype=2)+
  geom_segment(x=0,xend=150,y=1/6,yend=1/6,color="blue",linetype=2)+
  geom_segment(x=0,xend=150,y=1/4,yend=1/4,color="red",linetype=2) + 
  geom_point() + geom_line()
```

Write the tracks to a file

```{r}
write_csv(trackz_tbl,      file = "tracks_file.csv")
write_delim(trackz_tbl,    file = "tracks_file.txt",delim = "\t")
write_csv(trialz,  file = "headings_file.csv")
write_delim(trialz,file = "headings_file.txt",delim = "\t")
```


#### Radian conversion

```{r}
## change top of arena to be at zero and rotation clockwise
# trialz$abs_rads <- conversion.circular(
#                             trialz$abs_rads, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
# 
# trialz$radians <- conversion.circular(
#                             trialz$radians, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
# 
# # make the radians aligned to the top and rotation to be clockwise
# trackz_tbl$abs_theta <- conversion.circular(
#                             trackz_tbl$abs_theta, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
# 
# # make the radians aligned to the top and rotation to be clockwise
# trackz_tbl$rel_theta <- conversion.circular(
#                             trackz_tbl$rel_theta, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
```



#### Which videos are in spreadsheet but not present in tracks

Which videos in the df (Diadema tracks file are not represented in the tracks)

```{r}
### 
unique(df$video[which(!df$video %in% trialz$video)])
```
Which tracking files are not included in the spreadsheet.

```{r}
any(!trialz$video %in% df$video)
```






